---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  knitr::opts_chunk$set(error = TRUE),
  out.width="400px", dpi=400,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# CAFF

<!-- badges: start -->
<!-- badges: end -->

The goal of CAFF is to do Cell cycle Analysis of scRNA-seq data based on Fourier Filter.

## Installation

You can install the development version of `CAFF` like so:

``` r
# install.packages("devtools")
devtools::install_github("Xu_Feng/CAFF")
```

## Application
CAFF provides a variety of functions to analyze the cell cycle of single cell RNA-seq data in the following five aspects:
- Pseudotime quantification
- Identify the cell cycle-related genes
- Cell cycle phase labels
- Removing the cell cycle effect from the scRNA-seq dataset
- Determining G0 (quiescent) state

## Using the GSE129447(Hela) as an example:
### 1. Quality Control
#### 1.1 Cell and gene screening
- Gene screening: Firstly, the genes whose expression level exceeds 70% are 0 are removed. This percentage can fluctuate between 70-90% according to the quality of the data.
- Cell screening: Cells with sequencing depth less than 10000 were removed by viewing the sequencing depth distribution of the dataset.

```{r}
## Data preprocessing ------------------------
library(edgeR)
library(tidyverse)
library(CAFF)
library(data.table)

data(Hela1)

# Check the sequencing depth distribution:
quantile(colSums(Hela1), probs = c(0,0.1, 0.25,0.5,0.75,0.9, 1))

# Cells with sequencing depth less than 10000 were removed:
colids <- which(colSums(Hela1) > 10000)

# Genes with 0 expression in more than 70% of all cells were removed:
rowids <- which(apply(Hela1, 1, function(x) length(which(x == 0))/length(x) <= 0.7))

exp <- Hela1[rowids, colids]

```

#### 1.2 Data Preprocessing
- cpm transformation: For Count data, after quality control, the data is transformed to cpm format using the `cpm` function in the `edgeR` package.
- The expression matrix of cell cycle-related genes was extracted: the cell cycle genes used were 97 genes provided in Seurat package.

```{r}
# cpm transformation：
exp_cpm <- cpm(exp, log = T) %>% round(2)
exp_cpm_nolog <- cpm(exp, log = F) %>% round(2)

# extracte the expression matrix of cell cycle related genes:
load("cc_genes/cc_genes.Rdata")
cc_genes <- intersect(seurat_cc_genes, rownames(exp_cpm))
data_cc <- exp_cpm[cc_genes,]
```

#### 1.3 PCA was performed on data to observe data distribution and remove outliers

```{r fig.height=5, fig.width=8}
# the pca_plot function to see how the data is distributed on pc1 and pc2:
p_x <- pca_plot(data_cc, time.iqr = 1)
p_x$plot1
p_x$plot2

# remove_pc_outliers
final_index <- p_x$final_index
data_cc_new <- data_cc[, final_index]
```

### Pseudotime quantification
#### `CAFF` methods
```{r}
## Our method ==============
pseudo_order_list <- build_pseudo_order(data_cc_new, method = "Default")
pseudo_order <- pseudo_order_list$pseudo_order
pseudo_order_rank <- pseudo_order_list$pseudo_order_rank
```

#### `Tricycle` method
```{r}
## tricycle method ==============
tricycle_order_rank <- build_pseudo_order(data_cc_new, method = "tricycle")
```

#### `reCAT` method
```{r}
## reCAT method ==============
data_cc_cyclebase <-exp_cpm[intersect(cyclebase3.0_genes,rownames(exp_cpm)),intersect(colnames(data_cc_new),colnames(exp_cpm))]
dim(data_cc_cyclebase)


reCAT_order <- build_pseudo_order(data_cc_cyclebase, method ='reCAT')
reCAT_order_rank <- reCAT_order[[2]]
```

#### Comparison between the three methods
```{r fig.height=4, fig.width=12}
## Comparison between the three methods ==============
rank_data <- data.frame(pseudo_order_rank = pseudo_order_rank,
                         tricycle_order_rank = tricycle_order_rank,
                         reCAT_order_rank = reCAT_order_rank)
rank_data <- rank_data[order(rank_data$pseudo_order_rank),]
data_cc_ordered <- data_cc_new[, pseudo_order]

p_list <- cor_scatter_plot(rank_data)

library(cowplot)
p <- plot_grid(plotlist = p_list, ncol = 3)
p

```





